---
title: "Project Overview"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# written in R 3.6.2
library(tidyverse)
library(sf)
library(config)
library(pins)
library(pool)
library(DBI) # for dataset testing but use pool for any real work
library(lubridate)
```

## Background

Jason data request 12/3/2020:

"We're going to create two data sets (with Station ID, Lat, Long, median TN, median TP) by identifying stations from the last ten years where the median values of TN are less than 1 mg/L and TP less than 0.05 mg/L and a list of stations were median values of TN are over 2 mg/L or where the TP is over 0.1 mg/L.

Note the "low" nutrient stations need to have low TN and TP and the "high" nutrient station could be categorized as high by having a high TN or high TP. "


## Establish connection to ODS

```{r ODS connection}
# get configuration settings
conn <- config::get("connectionSettings")
# use API key to register board
board_register_rsconnect(key = conn$CONNECT_API_KEY,  #Sys.getenv("CONNECT_API_KEY"),
                          server = conn$CONNECT_SERVER)#Sys.getenv("CONNECT_SERVER"))
# establish db connection on server
pool <- dbPool(
  drv = odbc::odbc(),
  Driver = "SQL Server Native Client 11.0", 
  Server= "DEQ-SQLODS-PROD,50000",
  dbname = "ODS",
  trusted_connection = "yes"
)

```

## Start Small, Single Station Pull

### Station Data

```{r establish station}
station <- '2-JKS023.61'
```

Pull Station Field and location data. We need `Wqm_Field_Data_View$Fdt_Id` to get the appropriate analyte data from `Wqm_Analytes_View$Ana_Sam_Fdt_Id`.

```{r field data}
stationFieldData <- pool %>% tbl("Wqm_Field_Data_View") %>%
  filter(Fdt_Sta_Id %in% !! station &
           year(Fdt_Date_Time) >= year(!! Sys.Date()) - 10) %>% # limit to last 10 years of data on database side
  as_tibble()

stationGeospatial <- pool %>% tbl("WQM_Sta_GIS_View") %>%
  filter(Station_Id %in% !!station) %>%
  as_tibble()
```

There is a unique Fdt_Id for each sample event. We need these Fdt_Id's to filter the appropriate analyte data.

```{r analyte data}
stationAnalyteData <- pool %>% tbl("Wqm_Analytes_View") %>%
  filter(Ana_Sam_Fdt_Id %in% !! stationFieldData$Fdt_Id &
           year(Ana_Received_Date) >= year(!! Sys.Date()) - 10) %>% # limit to last 10 years of data on database side
  as_tibble() %>%
  left_join(dplyr::select(stationFieldData, Fdt_Id, Fdt_Sta_Id, Fdt_Date_Time), by = c("Ana_Sam_Fdt_Id" = "Fdt_Id")) 
```

Keep only the nutrient data we want.

```{r just nutrients}
stationNutrientData <-  stationAnalyteData %>%
    group_by(Ana_Sam_Fdt_Id, Fdt_Sta_Id, Fdt_Date_Time)%>%
    filter(Pg_Storet_Code == '00600' )%>%
    summarise(NITROGEN = sum(Ana_Value)) %>% # use conventionals new naming format
  left_join(stationAnalyteData %>%
              group_by(Ana_Sam_Fdt_Id, Fdt_Sta_Id, Fdt_Date_Time)%>%
              filter(Pg_Storet_Code == '00665' ) %>%
              summarise(PHOSPHORUS = sum(Ana_Value)), # use conventionals new naming format
            by = c('Ana_Sam_Fdt_Id','Fdt_Sta_Id', 'Fdt_Date_Time')) %>%
  ungroup()

```
Get summary statistics and join spatial information.

```{r stats and geospatial}
stationSummary <- stationNutrientData %>%
  group_by(Fdt_Sta_Id) %>%
  filter(!is.na(NITROGEN)) %>%
  summarise(medianTN = median(NITROGEN, na.rm = T),
            nTN = n()) %>%
  left_join(stationNutrientData %>% 
              group_by(Fdt_Sta_Id) %>%
              filter(!is.na(PHOSPHORUS)) %>%
              summarise(medianTP = median(PHOSPHORUS, na.rm = T),
                        nTP = n())) %>%
  left_join(dplyr::select(stationGeospatial, Station_Id, Latitude, Longitude, Deq_Region, County_City, Basin, Huc6_Vahu6, Huc6_Huc_8_Name),
            by = c('Fdt_Sta_Id' = 'Station_Id')) %>%
  st_as_sf(coords = c("Longitude", "Latitude"),  # make spatial layer using these columns
           remove = T, # remove these lat/lon cols from df
           crs = 4326) 

```



## Multistation Data Query

First we want to only pull info from stations that have analyte data in the last 10 years.

We need to break this up into multiple queries as to not blow up local memory or shut down ODS environment for everyone.

```{r nutrients in last year}
allNutrientData <- list()

for(i in year(Sys.Date()) - 0:1){#0:10){
  print(paste('Pulling data from : ', i))
  
  analyteData <- pool %>% tbl("Wqm_Analytes_View") %>%
    filter(year(Ana_Received_Date) >= !! i &
      #Ana_Sam_Fdt_Id %in% !! stationFieldData$Fdt_Id & # for testing
             Pg_Storet_Code %in% c('00600', '00665') ) %>% # limit to last 10 years of data on database side
    as_tibble()
  
  print(paste( i, ' Analyte Data complete'))
  
  nutrientData <- analyteData %>%
    group_by(Ana_Sam_Fdt_Id)%>%
    filter(Pg_Storet_Code == '00600' )%>%
    summarise(NITROGEN = sum(Ana_Value)) %>% # use conventionals new naming format
  left_join(analyteData %>%
              group_by(Ana_Sam_Fdt_Id)%>%
              filter(Pg_Storet_Code == '00665' ) %>%
              summarise(PHOSPHORUS = sum(Ana_Value)), # use conventionals new naming format
            by = c('Ana_Sam_Fdt_Id')) %>%
  ungroup()
  
  # Save data to list
  
  allNutrientData[[i]] <- list(nutrientData)
}

z <- purrr::map_df(allNutrientData, 1)


  
  # grab associated field data to get station name
  fieldData <- pool %>% tbl("Wqm_Field_Data_View") %>%
    filter(Fdt_Id %in% !! nutrientData$Ana_Sam_Fdt_Id ) %>%
    as_tibble() 
  
  print(paste( i, ' Field Data complete'))

  
  geospatial <- pool %>% tbl("WQM_Sta_GIS_View") %>%
    filter(Station_Id %in% !! unique(fieldData$Fdt_Sta_Id)) %>%
    as_tibble()
  
  print(paste( i, ' Geospatial Data complete'))
  
  nutrientData <- nutrientData %>%
  left_join(dplyr::select(fieldData, Fdt_Id, Fdt_Sta_Id, Fdt_Date_Time), by = c("Ana_Sam_Fdt_Id" = "Fdt_Id")) 

  # all together now
  yearlySummary <- nutrientData %>%
    group_by(Fdt_Sta_Id) %>%
    filter(!is.na(NITROGEN)) %>%
    summarise(medianTN = median(NITROGEN, na.rm = T),
              nTN = n()) %>%
    left_join(nutrientData %>% 
                group_by(Fdt_Sta_Id) %>%
                filter(!is.na(PHOSPHORUS)) %>%
                summarise(medianTP = median(PHOSPHORUS, na.rm = T),
                          nTP = n())) %>%
    left_join(dplyr::select(geospatial, Station_Id, Latitude, Longitude, Deq_Region, County_City, Basin, Huc6_Vahu6, Huc6_Huc_8_Name),
              by = c('Fdt_Sta_Id' = 'Station_Id'))
  
  # Save data to list
  
  allYearlySummary[[i]] <- list(yearlySummary)


}

```


```{r}
analyteData <- pool %>% tbl("Wqm_Analytes_View") %>%
  filter(year(Ana_Received_Date) >= year(!! Sys.Date()) - 0 &
    #Ana_Sam_Fdt_Id %in% !! stationFieldData$Fdt_Id & # for testing
           Pg_Storet_Code %in% c('00600', '00665') ) %>% # limit to last 10 years of data on database side
  as_tibble()

nutrientData <- analyteData %>%
    group_by(Ana_Sam_Fdt_Id)%>%
    filter(Pg_Storet_Code == '00600' )%>%
    summarise(NITROGEN = sum(Ana_Value)) %>% # use conventionals new naming format
  left_join(analyteData %>%
              group_by(Ana_Sam_Fdt_Id)%>%
              filter(Pg_Storet_Code == '00665' ) %>%
              summarise(PHOSPHORUS = sum(Ana_Value)), # use conventionals new naming format
            by = c('Ana_Sam_Fdt_Id')) %>%
  ungroup()

# grab associated field data to get station name
fieldData <- pool %>% tbl("Wqm_Field_Data_View") %>%
  filter(Fdt_Id %in% !! nutrientData$Ana_Sam_Fdt_Id ) %>%
  as_tibble() 

geospatial <- pool %>% tbl("WQM_Sta_GIS_View") %>%
  filter(Station_Id %in% !! unique(fieldData$Fdt_Sta_Id)) %>%
  as_tibble()

nutrientData <- nutrientData %>%
  left_join(dplyr::select(fieldData, Fdt_Id, Fdt_Sta_Id, Fdt_Date_Time), by = c("Ana_Sam_Fdt_Id" = "Fdt_Id")) 


# all together now
yearlySummary <- nutrientData %>%
  group_by(Fdt_Sta_Id) %>%
  filter(!is.na(NITROGEN)) %>%
  summarise(medianTN = median(NITROGEN, na.rm = T),
            nTN = n()) %>%
  left_join(nutrientData %>% 
              group_by(Fdt_Sta_Id) %>%
              filter(!is.na(PHOSPHORUS)) %>%
              summarise(medianTP = median(PHOSPHORUS, na.rm = T),
                        nTP = n())) %>%
  left_join(dplyr::select(geospatial, Station_Id, Latitude, Longitude, Deq_Region, County_City, Basin, Huc6_Vahu6, Huc6_Huc_8_Name),
            by = c('Fdt_Sta_Id' = 'Station_Id')) # %>%
  #st_as_sf(coords = c("Longitude", "Latitude"),  # make spatial layer using these columns
   #        remove = F, # dont remove these lat/lon cols from df
   #        crs = 4326) 


#allYearlySummary[[1]] <- list(yearlySummary)
#allYearlySummary[[2]] <- list(yearlySummary)

z <- purrr::map_df(allYearlySummary, 1)
```

